Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2021-11-18 manuscript/manuscript.tex
incorporate google docs feedback

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{Comparing human and model-based forecasts of COVID-19 in Germany and Poland}
\author{}
\date{\vspace{-2.5em}}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Comparing human and model-based forecasts of COVID-19 in Germany and Poland},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\emph{Anonymous Alpaca and many friends}

@@

Model-based forecasts, which have played an important role in shaping public policy throughout the COVID-19 pandemic, represent an implicit combination of model assumptions and the researcher's subjective opinion. This work analyses and compares human opinion against model-derived insights to discern relative strengths and weaknesses of both approaches. We compared purely opinion-derived forecasts of cases and deaths from COVID-19 in Germany and Poland, elicited from researchers and volunteers, against predictions from two semi-mechanistic epidemiological models. We also compared these forecasts against an ensemble of model-based, but expert-tuned, forecasts, submitted to the German and Polish Forecast Hub by other research institutions. In addition, we examined the effects of our contributions to the performance of the Hub ensemble. We found aggregated crowd forecasts to outperform all other methods when predicting cases (Weighted Interval Score relative to the Hub ensemble: 0.89), but not when predicting deaths (rel. WIS 1.26). Crowd forecasts were noticeably more overconfident (55\% and 75\% coverage of the 90\% prediction intervals for cases and deaths, respectively) than model-based predictions. Performance of the semi-mechanistic models was good short-term, but deteriorated quickly over time when assumptions were no longer met.

@@

Infectious disease forecasting has a long tradition and has helped inform public health decisions about (among others) influenza (e.g. McGowan et al. 2019; Reich et al. 2019; Shaman and Karspeck 2012; Biggerstaff et al. 2016), dengue fever (e.g. Johansson et al. 2019; Yamana, Kandula, and Shaman 2016; Colón-González et al. 2021), ebola (e.g. Viboud et al. 2018; Funk et al. 2019), chikungunya (e.g. Del Valle et al. 2018; Farrow et al. 2017) and now COVID-19 (e.g. Funk et al. 2020; Cramer et al. 2020; Cramer et al. 2021; Bracher, Wolffram, et al. 2021; Hub 2021). The aggregation of forecasts of COVID-19 cases, hospitalisations or deaths from multiple teams has created opportunities to compare the predictive performance of different models. Yet it has been difficult to discern which factors lead to good COVID-19 forecasts. Answering this question is complicated by the varying and generally unobserved degree of manual expert intervention. Model-based forecasts usually represent an implicit combination of human judgement and model-derived insights, in the sense that forecasts have been generated with the help of epidemiological or statistical models, but throughout the process were manually changed and adjusted according to the researchers' judgement.

Previous work has examined human predictions in various contexts, such as geopolitics (Tetlock et al. 2014; Atanasov et al. 2016), meta-science (Hoogeveen, Sarafoglou, and Wagenmakers 2020; ReplicationMarkets 2020), sports (Servan-Schreiber et al. 2004) and epidemiology (Farrow et al. 2017; McAndrew and Reich 2020; Recchia, Freeman, and Spiegelhalter 2021). Several prediction platforms (Metaculus 2020; Hypermind 2021; CSET Foretell 2021) and prediction markets (PredictIt 2021) have been created to collate expert and non-expert predictions. However, with the notable exception of Farrow et al., these crowd forecasts were not designed to be evaluated alongside model-based forecasts and usually follow their own (often binary) formats. Farrow et al.~found that humans could outperform computer models at predicting the 2014/15 and 2015/16 flu season in the US, a setting where the disease was well known and information about previous seasons was available. They tended to do slightly worse at predicting the 2014/15 outbreak of chikungunya in the Americas, a disease previously largely unobserved and unknown there and then.

In this study, we attempt to further shed light on the role of human judgement and model-derived insight and analyse respective strengths and weaknesses. To allow for a direct comparison on an equal footing we employ a novel approach that elicits full probabilistic forecasts (Held, Meyer, and Bracher 2017) from participants, rather than constructing a predictive distribution from individual point predictions.

@@

We created a crowd-sourced forecast (``crowd forecast''), elicited from participants through a web application (Bosse et al. 2020b), as well as two semi-mechanistic models (``renewal model'' and ``convolution model'') informed by basic assumptions about COVID-19 epidemiology. Forecasts were created in real time over a period of 21 weeks from October 12th 2020 until March 1st 2021 and submitted to the German and Polish Forecast hub (Bracher, Wolffram, et al. 2021). All code and tools necessary to generate the forecasts and make a forecast submission are available in the \texttt{covid.german.forecasts} R package (Bosse et al. 2020a). This repository also contains a record of all forecasts submitted to the German and Polish Forecast Hub. Forecasts were evaluated using a variety of scoring metrics and compared among each other and against an ensemble of all other models submitted to the Forecast Hub.

@@

The German and Polish Forecast Hub elicits predictions for various COVID-19 related forecast targets from different research institutions every week. Forecasts had to be made every Monday (with submissions allowed until Tuesday 3pm) using data available until Monday 11.59pm. We submitted forecasts for incident and cumulative weekly reported numbers of cases and deaths from COVID-19 on a national level in Germany and Poland over a one to four week forecast horizon. Weeks were defined as epiweeks starting on Sunday and ending on Saturday, meaning that forecast horizons were in fact 5, 12, 19 and 26 days. Submissions were required in a quantile-based format with 23 quantiles at levels 0.01, 0.025, 0.05, 0.10, 0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub were combined to different ensembles every week, with the median ensemble being the default ensemble shown on all official Forecast hub visualisations.

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish forecast hub. Until December 14th 2020 these data were sourced from the European Centre for Disease Control (ECDC) (ECDC 2020). After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period (RKI 2021). These data are subject to reporting artefacts (such as a delayed case reporting in Poland on the 24th November (Forsal.pl 2020)), changes in reporting over time and variation in testing regimes (e.g.~in Germany from the 11th of November on (Ärzteblatt 2020)).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (\url{https://cmmid-lshtm.shinyapps.io/crowd-forecast/}). Forecasts were allowed until 12pm on Tuesdays, but participants were asked to only use information available until Monday. The application was built using the shiny and golem R packages (Fay et al. 2021; Chang et al. 2021) and is available in the \texttt{crowdforecastr} R package (Bosse et al. 2020b). To make a forecast in the application participants could select a predictive distribution, with the default being log-normal, and adjust the median and the width of the uncertainty by either interacting with a figure showing their forecast or providing numerical values. The baseline shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in log observed forecasts (i.e.~as \(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
Our interface also allowed participants to view the observed data, and their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from ({``{COVID}-19 {Data Explorer}''} n.d.). These data included e.g.~notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time.

Forecasts were stored in a Google Sheet and downloaded, cleaned and processed every week for submission. If a forecaster had submitted multiple predictions for a single target, only the latest submission was kept. Information on the chosen distribution as well as the parameters for median and width were used to obtain the required set of 23 quantiles from that distribution. Forecasts from all forecasters were then aggregated using an unweighted quantile-wise mean. On a few occasions, individual forecasts were assessed as clearly erroneous by visual inspection and subsequently removed before aggregation.

Participants were recruited mostly within the Centre of Mathematical Modeling of Infectious Diseases at the London School of Hygiene \& Tropical Medicine, but participants were also invited personally or via social media to submit predictions. Depending on whether they had a background in either statistics, forecasting or epidemiology, participants were asked to self-identify as `experts' or `non-experts.'

@@

We used two models from the \texttt{EpiNow2} R package (version 1.3.3) as our model-based forecasts (Abbott, Hellewell, et al. 2020). The first of these models, here called ``renewal model,'' used the renewal equation (Fraser 2007) to predict reported cases and deaths. It estimated the effective reproduction number \(R_t\) (the average number of people each infected person is expected to infect in turn) and models future infections as a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant beyond the forecast horizon, roughly corresponding to continuing the latest exponential trend in infections. On the 9th of November we altered the date when \(R_t\) was assumed to be constant from two weeks prior to the date of the forecast to the forecast date. Reported notifications were obtained by convolving predicted infections over data-based delay distributions (Abbott, Hellewell, et al. 2020; epiforecasts.io/covid n.d.; Sherratt et al. 2021) to model the time between infection and report date or death. The renewal model was used to predict cases as well as deaths with forecasts being generated for each target separately, meaning that \(R_t\) was forecast once from reported cases and once from reported deaths. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was delayed and therefore not included in the official Forecast hub ensemble.

The second model (``convolution model'') was only used to forecast deaths and was added later, starting December 7th 2020 (with the first forecast from December 7th suffering from a software bug and therefore disregarded in all further analyses). The convolution model was submitted, but never included in the official Forecast hub ensemble due to concerns that it could be too similar to the renewal model. The convolution model predicted deaths as a fraction of infected people who would die with some delay, by using a convolution of reported cases with a distribution that described the delay from case report to death and a scaling factor (the case-fatality ratio). Both the renewal and the convolution model used daily observations and assumed a negative binomial observation model with a multiplicative day-of-the-week effect (Abbott, Hellewell, et al. 2020). Line list data used to inform the prior for the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from (Xu et al. 2020) with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team 2020) with each location and forecast target being fit separately. More details are available in the supplementary information.

@@

For the main analysis we focused on two week ahead predictions, as predictions beyond this horizon are often unreliable due to rapidly changing conditions (Bracher, Wolffram, et al. 2021). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period where forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Comparing human and model-based forecasts of COVID-19 in Germany and Poland},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Comparing human and model-based forecasts of COVID-19 in Germany and Poland}
\author{}
\date{\vspace{-2.5em}}

@@

Target journal: eLife

\emph{Nikos I. Bosse, Sam Abbott, Johannes Bracher, Habakuk Hain, Billy J. Quilty, Mark Jit, CMMID Working Group, Edwin van Leeuwen, Anne Cori, Sebastian Funk}

@@

Forecasts based on epidemiological modelling have played an important role in shaping public policy throughout the COVID-19 pandemic. This modelling combines knowledge about infectious disease dynamics with the subjective opinion of the researcher who develops and refines the model and often also adjusts model outputs. Developing a forecast model is difficult, resource- and time-consuming. It is therefore worth asking what modelling is able to add beyond the subjective opinion of the researcher alone. To investigate this, we analysed different real-time forecasts of cases of and deaths from COVID-19 in Germany and Poland over a 1-4 week horizon submitted to the German and Polish Forecast Hub. We compared crowd forecasts elicited from researchers and volunteers, against a) forecasts from two semi-mechanistic models based on common epidemiological assumptions and b) the ensemble of all other models submitted to the Forecast Hub. We found crowd forecasts, despite being overconfident, to outperform all other methods across all forecast horizons when forecasting cases (weighted interval score relative to the Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational models performed comparably better when predicting deaths (rel. WIS 1.26), suggesting that epidemiological modelling and human judgement can complement each other in important ways.

@@

Infectious disease modelling has a long tradition and has helped inform public health decisions both through scenario modelling, as well as actual forecasts of (among others) influenza (e.g. McGowan et al. 2019; Reich et al. 2019; Shaman and Karspeck 2012; Biggerstaff et al. 2016), dengue fever (e.g. Johansson et al. 2019; Yamana, Kandula, and Shaman 2016; Colón-González et al. 2021), ebola (e.g. Viboud et al. 2018; Funk et al. 2019), chikungunya (e.g. Del Valle et al. 2018; Farrow et al. 2017) and now COVID-19 (e.g. Funk et al. 2020; Cramer et al. 2020; Cramer et al. 2021; Johannes Bracher, Wolffram, et al. 2021; J. Bracher et al. 2021; Hub 2021). Applications of epidemiological models differ in the way they make statements about the future. Forecasts aim to predict the future as it will occur, while scenario modelling and projections aim to represent what the future could look like under certain scenario assumptions or if conditions stayed the same as they were in the past. Forecasts can be judged by comparing them against observed data. Since it is much harder to fairly assess the accuracy and usefulness of projections and scenario modelling in the same way, this work focuses on forecasts, which represent only a subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been collected, aggregated and compared by Forecast Hubs such as the US Forecast Hub (Cramer et al. 2020; Cramer et al. 2021), the German and Polish Forecast Hub (Johannes Bracher, Wolffram, et al. 2021; J. Bracher et al. 2021) and the European Forecast Hub (Hub 2021). Often, different individual forecasts are combined into a single forecast, e.g.~by taking the mean or median of all forecasts. These ensemble forecasts usually tend to perform better and more consistently than individual forecasts {[}see e.g. Yamana, Kandula, and Shaman (2016); reichAccuracyRealtimeMultimodel2019{]}.

Individual computational models usually rely to varying degrees on mechanistic assumptions about infectious disease dynamics (such as SIR-type compartmental models that aim to represent how individuals move from being susceptible to infected and then recovered or dead) or are more statistical in nature (such as time series models that detect statistical patterns without explicitly modelling disease dynamics). How exactly such a mathematical or computational model is constructed and which assumptions are made depends on subjective opinion and judgement of the researcher who develops and refines the model. Models are commonly adjusted and improved based on whether the model output looks plausible to the researchers involved.

The process of model construction and refinement is laborious and time-consuming, and it is therefore worth asking what modelling can add beyond the subjective judgment of the researcher alone. In this work, we ask this question specifically in the context of predictive performance, and set aside other advantages of epidemiological modelling (such as reproducibility or the ability to obtain a deeper fundamental understanding of how diseases spread). One natural way to do this is to compare the predictive performance of forecasts based on computational models (``model-based forecasts'') against forecasts made by individual humans without explicit use of a computer model (``direct human forecasts'') or a combination of multiple such forecasts (``crowd forecasts'').

Previous work has examined such direct human forecasts in various contexts, such as geopolitics (Tetlock et al. 2014; Atanasov et al. 2016), meta-science (Hoogeveen, Sarafoglou, and Wagenmakers 2020; ReplicationMarkets 2020), sports (Servan-Schreiber et al. 2004) and epidemiology (Farrow et al. 2017; McAndrew and Reich 2020; Recchia, Freeman, and Spiegelhalter 2021). Several prediction platforms (Metaculus 2020; Hypermind 2021; CSET Foretell 2021) and prediction markets (PredictIt 2021) have been created to collate expert and non-expert predictions. However, with the notable exception of Farrow et al. (2017), these forecasts were not designed to be evaluated alongside model-based forecasts and usually follow their own (often binary) prediction formats. Direct human forecasts may be able to take into account insights and relationships between variables which are hard to specify using epidemiological models. However, it is not entirely clear in which situations human forecasts perform well or badly. For example, Farrow et al. (2017) found that humans could outperform computer models at predicting the 2014/15 and 2015/16 flu season in the US, a setting where the disease was well known and information about previous seasons was available. However, humans tended to do slightly worse at predicting the 2014/15 outbreak of chikungunya in the Americas, a disease previously largely unobserved and unknown there and then.

In this study, we analyse the performance of direct human forecasts relative to model-based forecasts and discuss the added benefit of epidemiological modelling over human judgement alone. As a case study, we use different forecasts, involving varying degrees of human intervention, which we submitted in real time to the German and Polish Forecast Hub. In contrast to Farrow et al. (2017) we elicited not only point predictions, but full predictive distributions (``probabilistic forecasts'') (Held, Meyer, and Bracher 2017) from participants. This allows us to compare not only predictive accuracy, but also how well human forecasters and model-based forecasts were able to quantify the appropriate forecast uncertainty.

@@

We created and submitted the following forecasts to the German and Polish Forecast Hub: 1) a direct human forecast (henceforth called ``crowd forecast''), elicited from participants through a web application (N. I. Bosse et al. 2020b) and 2) two semi-mechanistic model-based forecasts (``renewal model'' and ``convolution model'') informed by basic assumptions about COVID-19 epidemiology. While the two semi-mechanistic forecasts were necessarily shaped by our implicit assumptions and decisions, they were designed such as to minimise the amount of human intervention involved. For example, we refrained from adjusting model outputs or refining the models based on past performance. Forecasts were created in real time over a period of 21 weeks from October 12th 2020 until March 1st 2021 and submitted to the German and Polish Forecast hub (Johannes Bracher, Wolffram, et al. 2021; J. Bracher et al. 2021). All code and tools necessary to generate the forecasts and make a forecast submission are available in the \texttt{covid.german.forecasts} R package (N. I. Bosse et al. 2020a). This repository also contains a record of all forecasts submitted to the German and Polish Forecast Hub. Forecasts were evaluated using a variety of scoring metrics and compared among each other and against an ensemble of all other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub elicits predictions for various COVID-19 related forecast targets from different research institutions every week. Forecasts had to be made every Monday (with submissions allowed until Tuesday 3pm) and were permitted to use any data that was available by Monday 11.59pm. We submitted forecasts for incident and cumulative weekly reported numbers of cases of and deaths from COVID-19 on a national level in Germany and Poland over a one to four week forecast horizon. Forecasts were submitted on Mondays, but weeks were defined as ending on a Saturday (and starting on Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26 days. Submissions were required in a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub were combined into different ensembles every week, with the median ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the median of all submitted \(\alpha\)-quantiles) being the default ensemble shown on all official Forecast hub visualisations (\url{https://kitmetricslab.github.io/forecasthub/forecast}).

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish forecast hub. Until December 14th, 2020, these data were sourced from the European Centre for Disease Control (ECDC) (ECDC 2020). After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period (RKI 2021). These data are subject to reporting artefacts (such as a delayed case reporting in Poland on the 24th November (Forsal.pl 2020)), changes in reporting over time and variation in testing regimes (e.g.~in Germany from the 11th of November on (Ärzteblatt 2020)).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (\url{https://cmmid-lshtm.shinyapps.io/crowd-forecast/}). Weekly forecasts had to be submitted before Tuesday 12pm every week, but participants were asked to only use any information or data that was already available by Monday night. The application was built using the shiny and golem R packages (Fay et al. 2021; Chang et al. 2021) and is available in the \texttt{crowdforecastr} R package (N. I. Bosse et al. 2020b). To make a forecast in the application participants could select a predictive distribution representing the probability that the forecasted quantity took certain values, with the default being log-normal, and adjust the median and the width of the uncertainty by either interacting with a figure showing their forecast or providing numerical values. {[}SCREENSHOT OF APPLICATION IN SI{]} The baseline shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e.~as \(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
Our interface also allowed participants to view past observations based on the hub data, as well as their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from (Data 2020). These data included, for example, notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time.

Forecasts were stored in a Google Sheet and downloaded, cleaned and processed every week for submission to the Forecast Hub. If a forecaster had submitted multiple predictions for a single target, only the latest submission was kept. Information on the chosen distribution as well as the parameters for median and width were used to obtain the required set of 23 quantiles from that distribution. Forecasts from all forecasters were then aggregated using an unweighted quantile-wise mean (i.e., the \(\alpha\)-quantile of the ensemble is given by the mean of all submitted \(\alpha\)-quantiles). On a few occasions, individual forecasts were assessed as clearly erroneous by visual inspection and subsequently removed before aggregation and were excluded from the submission as well as the analysis.

Participants were recruited mostly within the Centre of Mathematical Modeling of Infectious Diseases at the London School of Hygiene \& Tropical Medicine, but participants were also invited personally or via social media to submit predictions. Depending on whether they had a background in either statistics, forecasting or epidemiology, participants were asked to self-identify as `experts' or `non-experts'.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R package (version 1.3.3) as our model-based forecasts (Abbott, Hellewell, Hickson, et al. 2020). The first of these models, here called ``renewal model'', used the renewal equation (Fraser 2007) to predict reported cases and deaths (see details in the SI). It estimated the effective reproduction number \(R_t\) (the average number of people each person infected at time t is expected to infect in turn) and modelled future infections as a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant beyond the forecast date, roughly corresponding to continuing the latest exponential trend in infections. On the 9th of November we altered the date when \(R_t\) was assumed to be constant from two weeks prior to the date of the forecast to the forecast date, which we found to yield a more stable \(R_t\) estimate. Reported case and death notifications were obtained by convolving predicted infections over data-based delay distributions (Abbott, Hellewell, Hickson, et al. 2020; epiforecasts.io/covid n.d.; Sherratt et al. 2021; Abbott, Hellewell, Thompson, et al. 2020) to model the time between infection and report date. The renewal model was used to predict cases as well as deaths with forecasts being generated for each target separately. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was delayed and therefore not included in the official Forecast hub ensemble.

The second model (``convolution model'', see details in SI) was only used to forecast deaths and was added later, starting December 7th 2020 (with the first forecast from December 7th suffering from a software bug and therefore disregarded in all further analyses). The convolution model was submitted, but never included in the official Forecast hub ensemble due to concerns that it could be too similar to the renewal model. The convolution model predicted deaths as a fraction of infected people who would die with some delay, by using a convolution of reported cases with a distribution that described the delay from case report to death and a scaling factor (the case-fatality ratio). Both the renewal and the convolution model used daily observations and assumed a negative binomial observation model with a multiplicative day-of-the-week effect (Abbott, Hellewell, Hickson, et al. 2020).

Line list data used to inform the prior for the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from (Xu et al. 2020) with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team 2020) with each location and forecast target being fitted separately.

@@

For the main analysis we focused mostly on two week ahead forecasts, as COVID-19 forecasts, especially for cases, were in the past found to have poor predictive performance beyond this horizon (Johannes Bracher, Wolffram, et al. 2021). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period for which all forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) (Johannes Bracher, Ray, et al. 2021), the absolute error, relative bias, and empirical coverage of the 50\% and 90\% prediction intervals. The WIS is a proper scoring rule (Gneiting and Raftery 2007), meaning that in expectation the score is optimised by reporting a predictive distribution that is identical to the true data-generating distribution. Forecasters are therefore incentivised to report their true belief about the future. The WIS can be understood as a generalisation of the absolute error to quantile-based forecasts (also meaning that smaller values are better) and can be decomposed into three separate penalties: forecast spread (i.e.~uncertainty of forecasts), over-prediction and under-prediction. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric (Funk et al. 2019) we use captures how much probability mass of the forecast was above or below the true value (mapped to values between -1 and 1) and therefore represents a general tendency to over- or under-predict in relative terms. A value of -1 implies that all quantiles of the predictive distribution are below the observed value and a value of 1 that all quantiles are above the observed value. Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g.~how many observed values fall inside all 50\% prediction intervals). Scoring metrics are explained in more detail in Table \ref{tab:scoring-metrics} in the SI. All scores were calculated using the \texttt{scoringutils} R package (N. Bosse 2020).
